---
title: "Assignment 5 Prediction"
author: "Zhulin Yu"
date: "11/14/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries

```{r Load Libraries}
library(rpart)
library(party)
```

Load Data

```{r Load Data}
D1 = read.csv("intelligent_tutor.csv")
```

Here are all the variables:

id - student id prior_prob_count - The number of problems a student has done in the system prior to the surrent session

score - The score the student achieved in the current session

hints - The number of hints the student requested in the current session

hint.y - Whether or not the student asked for hints in the current session

complete - Whether or not the student completed the current session

prior_prob_count - total number of problems solved in previous sessions

prior_percent_correct - percentage of correct in previous sessions

action - The action suggested by the system to a teacher about a given student based on their performance: 1 = Ask Teacher; 2 = Start New Session; 3 = Give up.

# Part I, Build a classification tree to predict behavior

```{r }
c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)

#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")
```

# Part II, Regression Tree

Build a regression tree to predict students' final scores, which determine what tpye of intervention is preferred.

First, Visulaize score data to decide "leaves"
```{r}
hist(D1$score)
```

Define "Leaves": score <= 0.4, intervene; score between 0.4 and 0.8, monitor; score >= 0.8, no action.

Then generate the regression tree by using ctree. And plot.

```{r}
D1$advice <- ifelse(D1$score <=0.4, "intervene", ifelse(D1$score > 0.4 & D1$score <=0.8, "monitor", "no action"))
score_ctree <- ctree(factor(advice) ~ prior_prob_count + prior_percent_correct + hints, D1)
plot(score_ctree)
```

As shown in the figure above, intervention decisions would be made in the following procedure:

1, If one has asked for hints:

    1.1, And he/she asked for more than 12 hints, we do not need to take any actions.

    1.2, But he/she asked for 12 hints or less, we will check his/her preiovus precentage of correctness:

        1.2.1, If the percentage is 0.629 or less, we might want to moniter him/her or even intervene.

        1.2.2, If the percentage is higher than 0.629, we might want to moniter him/her.

        * However, p = 0.047 and it indicates that the difference between the two groups, if only considering the mean, is marginally             significant. Therefore, it could be quite objective to make the decision.

2, If he/she never asked for hints:

    2.1, And he/she has solved 85 or less problems before, we might want to monitor or take no action.

    2.2, And he/she has solved more than 85 problems, we would take no action.
    
To conclude, I would recommend to pay more attention to number of hints (hints) and total number of problems solved before (prior_prob_count).

### Using the regression tree to predict upon new data
```{r}
# import new dataset
D2 = read.csv("intelligent_tutor_new.csv")

#Generate predicted advice for new students based on tree generated from old students
D2$prediction <- predict(score_ctree, D2)
```

# Part III
Compare the predicted advice with the actual advice that these studnts received. What is the difference between the observed and predicted results?

```{r}
# Recode "no action" to 3; "monitor" to 2; and "intervene" to 1.
D2$prediction_num[D2$prediction == "no action"] = 3
D2$prediction_num[D2$prediction == "monitor"] = 2
D2$prediction_num[D2$prediction == "intervene"] = 1

# Then visualize the distribution of the three actions.
hist(D2$prediction_num)
```

In the observed result, all student have score of 1, thus they received "no action". However, in the predicted result, many have "monitor" (80 out of 200). The remaining students are predicted correctly. The accuracy is 60%.